Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-28 paper/high-assurance-contracts.tex
Minor upgrades

\subsubsection*{Imperative code contracts}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages targeted by SPLAT do not directly support streams.  We are
developing a translation to eliminate such deep temporal accesses by
the introduction of intermediate variables. The essential property the
translation needs to satisfy is a syntactic restriction ---call
it \emph{Pascalish}---that supports easy translation to standard
imperative languages.  The guiding intuition is that a Pascalish
{\Eqs} accesses at most one step in the past; also, accesses to a
past value of a variable are only possible until the variable is
updated. In informal terms, this means that ${\Eqs}\cdot G$ in a code
contract has the following properties:
\begin{itemize}
\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-Pascalish manner, by
{\small
\begin{lstlisting}[style=agree]
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{lstlisting}
}
\noindent or by (again non-Pascalish)
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
\end{lstlisting}
}
\noindent or by (Pascalish):
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  F2 = 42 -> pre(F1)
  F1 = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else F1 + F2
\end{lstlisting}
} A Pascalish ${\Eqs} \cdot G$ is straightforward to implement in
conventional programming languages, as we will see in
Section \ref{sec:synthesis}.

>>>>

\subsubsection*{Imperative code contracts}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages targeted by SPLAT do not directly support streams.  We are
developing a translation to eliminate such deep temporal accesses by
the introduction of intermediate variables. The essential property the
translation needs to satisfy is a syntactic restriction ---call
it \emph{Pascalish}---that supports easy translation to standard
imperative languages.  The guiding intuition is that a Pascalish
{\Eqs} accesses at most one step in the past; also, accesses to a
past value of a variable are only possible until the variable is
updated. In informal terms, this means that ${\Eqs}\cdot G$ in a code
contract has the following properties:
\begin{itemize}
\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-Pascalish manner, by
{\small
\begin{lstlisting}[style=agree]
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{lstlisting}
}
\noindent or by (again non-Pascalish)
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
\end{lstlisting}
}
\noindent or by (Pascalish):
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  F2 = 42 -> pre(F1)
  F1 = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else F1 + F2
\end{lstlisting}
}
(Note that the initial values of \verb+F1+ and \verb+F2+ are never
accessed, so 42 could be replaced by any number in their defining
equations.) A Pascalish ${\Eqs} \cdot G$ is straightforward to
implement in conventional programming languages, as we will see in
Section \ref{sec:synthesis}.

### 2022-03-28 paper/verification-conditions.tex
Minor upgrades

\subsection{Components}
A system implementation is described by its subcomponents plus
connections between their ports.  When considering a system of
components, data flows through the components in dependency order,
with inputs being propagated to outputs through all contracts until
they stabilize (i.e., they cannot propagate further). Therefore, the
subcomponent contracts, and thus the top-level model, must be acyclic.
Cyclic systems are made acyclic by breaking cycles with delays.  Once
the data propagation has stabilized, the model proceeds to the next
input data in the input streams. It is worth noting that the semantics
do not model computation or communication delay. The output of one
contract is seen at the input of any downstream contract in the same
step of the input data stream.

An component in AGREE is described relative to its \emph{ports}, a
concept inherited from AADL.  There are three types of AADL ports:
\emph{event} ports, \emph{data} ports, and \emph{event data} ports.
An event port has no data and only signals occurrence (or not) of the
event. A data port is a sampled port: the data can be read at anytime.
An event data port indicates when data is present with an event.

@@

If all the verification conditions hold (AGREE uses $k$-inductive
model checking to automatically prove or disprove each generated
verification condition), then the system is said to be \emph{correct},
meaning that the system composition meets input assumptions at each
input as well as the guarantees on the system output. A consequence of
this result is that $\globally(\historically(\assumes) \implies
\guarantees)$ holds for the system contract.
Therefore the Liskov substitution principle applies, and as such, the implementation is a safe substitution for the system contract \cite{10.1145/62139.62141}.

The expanded property lists in \figref{fig:example-certificate} and
\figref{fig:hardened-certificate} are the results from verifying or
disproving the above verification conditions.  The additional
unexpanded results at the bottom of the figures prove
\emph{self-consistency} in the contracts.  It is not uncommon to
accidentally write contracts that are self-contradicting.  For
example, a contract may guarantee an output be two different values in
the same moment of time.  AGREE generates additional verification
conditions that prove each component contract, and the composition of
contracts, self-consistent.

>>>>

\subsection{Components}
A system implementation is described by its ports, its subcomponents,
and the connections between ports on different subcomponents.  Ports
in AGREE are inherited from AADL.  There are three kinds of port:
\emph{event} ports, \emph{data} ports, and \emph{event data} ports.
An event port carries no data and only signals occurrence (or not) of the
event. A data port is a sampled port: the data can be read at anytime.
An event data port indicates when data is present with an event.

Data flows through the subcomponents of a system in dependency order,
with inputs being propagated to outputs through all contracts until
they stabilize \ie, cannot propagate further. Therefore, subcomponent
contracts, and thus the top-level model, must be acyclic.  Cyclic
systems are made acyclic by breaking cycles with delays.  Once the
data propagation has stabilized, the model proceeds to the next input
data in the input streams. It is worth noting that the semantics do
not model computation or communication delay: the output of one
contract is seen at the input of any downstream contract in the same
step of the input data stream.

@@

If all the verification conditions hold (AGREE uses $k$-inductive
model checking to automatically prove or disprove each generated
verification condition), then the system is said to be \emph{correct},
meaning that the system composition meets input assumptions at each
input as well as the guarantees on the system output. A consequence of
this result is that $\globally(\historically(\assumes) \implies
\guarantees)$ holds for the system contract.  Therefore the Liskov
substitution principle applies and the implementation is a safe
substitution for the system contract \cite{10.1145/62139.62141}.

The expanded property lists in \figref{fig:example-certificate} and
\figref{fig:hardened-certificate} are the results from verifying or
disproving the above verification conditions.  The additional
unexpanded results at the bottom of the figures prove
\emph{self-consistency} in the contracts.\emph{Needs a definition?}  It is not uncommon to
accidentally write contracts that are self-contradicting.  For
example, a contract may guarantee an output be two different values in
the same moment of time.  AGREE generates additional verification
conditions that prove each component contract, and the composition of
contracts, self-consistent.
